---
title: "Equality of Opportunity Seminar Paper"
author: "Leonard Fidlin, Daniel Jost, Anne Valder"
date: "14 1 2021"
bibliography: bibliography.bib
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: false
subtitle: Data Science and Machine Learning 2187 & 2087
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(fig.align = 'center')
```

# **Introduction**
```{r echo=T, message=FALSE, error=FALSE, warning=FALSE}
library(rpart)
library(rpart.plot) 
# library(precrec)
library(caret)       
library(party)
# library(partykit)
library(tidyverse)
# library(vcd)
library(RColorBrewer)
library(knitr)
# library(glmnet)  
# library(haven)
# library(fst)
# library(ranger)
# library(tuneRanger)
# library(xgboost)
library(readr)
library(summarytools)
```
# **Data Wrangling**
We use the EU-SILC data from 2011 as it was the survey when additionally there were questions on inter-generational transmission. These were questions on the parents of the respondents. We want to see, whether it is possible using only the information given about the parents to predict the correct income of the respondents and what the outstanding circumstances are for the trees we grow.

The unique identifier used in all  four data sets is the household ID identifier: RX030 in the Personal Register, PX030 in the Personal Data, DB030 in the Household Register, and HB030 in the Household Data file.

We use the following variables for circumstances: Respondent's sex (PB150), Respondent's country of birth (Citizenship? PB220A), Presence of parents at home (PT010), Number of adults (18 or older) in respondents household (PT020), Number of working adults (18 or older) in respondents household (PT030), Father/Mother country of birth and citizenship (PT060, PT070, PT090, PT100), Father/mother education (PT110, PT120), Father/mother occupational status (PT130, PT160), Father/mother main occupation (PT150,PT180), Managerial position of father/mother(PT140,PT170), Tenancy status of the house in which respondent was living as a child (PT210).

Outcome Variables i.e. Income: Total Household gross income (HY010), Total Disposable Income (HY020), Dwelling Type (HH010), Housing (HH030).

```{r echo=T, message=FALSE, error=FALSE, warning=FALSE}
# setting the data path
data_path ="./AT2011"
getwd()

# accessing the data
personal_data <- read.csv(file.path(data_path, "AT_2011p_EUSILC.csv"))
household_data <- read.csv(file.path(data_path, "AT_2011h_EUSILC.csv"))

# change the name of the identifier variable
household_data <- household_data %>% rename("PX030" = HB030)

# joining the data
equality_data <- personal_data %>%  left_join(household_data, by = "PX030")

# Renaming important variables for readability of tree
equality_data <- equality_data %>% mutate(
  
  log_income = log(HY010 + 3)) %>% mutate( #We add 3 to the Income variable to get rid of negative values, we assume that this does not significantly impact the distribution of the variable over all. 
    # This is questionable, we should remove those with negative answers instead. And do the Winsorize Method to clean the Income Data
    Age = (2011 - PB140)
  ) %>% #Age Is currently Missing
  rename("Sex" = PB150,
         "Citizenship" = PB220A,
         "Parents_Present" = PT010,
         "Adults_home" = PT020, 
         "Working_Adults" = PT030,
         "Father_CoB" = PT060,
         "Father_Cit" = PT070,
         "Mother_CoB" = PT090,
         "Mother_Cit" = PT100,
         "Father_Educ" = PT110,
         "Mother_Educ" = PT120,
         "Father_Occup_Stat" = PT130,
         "Mother_Occup_Stat" = PT160,
         "Father_Main_Occup" = PT150,
         "Mother_Main_Occup" = PT180,
         "Father_Manag" = PT140,
         "Mother_Manag" = PT170,
         "Tenancy" = PT210,)

equality_data_1 <- equality_data %>% select( # Age is missing!
  Sex, Age, Citizenship, Parents_Present, Adults_home, Working_Adults, Father_CoB, Father_Cit, Mother_CoB, Mother_Cit, Father_Educ, Mother_Educ, Father_Occup_Stat, Mother_Occup_Stat, Father_Main_Occup, Mother_Main_Occup, Father_Manag, Mother_Manag, Tenancy, log_income
)


print(dfSummary(equality_data_1), method="render")

```
# **Method**: Conditional Inference Trees

* `ctree` from party package in R
* recursive partitioning just like `rpart`
* `rpart`: maximizing an information measure
* `ctree`: significance test procedure

### Advantages

**Advantages of Trees:** straightforward to interpret

**Advantages of Trees over linear regression models:** very large set of observations can be used & model speciﬁcation is no longer exogenously given

**Advantages of Conditional Inference Trees over Regression and Classification Trees (CART):** the algorithm automatically provides a test for the null hypothesis of equality of opportunity & prevents overfitting while CART "cannot distinguish between a significant and an insignificant improvement in the information measure" (Mingers 1987, as cited in @hot, 2) & consider the distributional properties of the measures.


### Procedure 

The algorithm follows a stepwise procedure [@brunori20, 7-8]:

1. **Choose confidence level**
2. **Choose feature:** test all the null hypotheses of independence between the individual outcome and each of all the observable circumstances
    i) no hypothesis can be rejected: stop
    ii) one or more circumstance is siginificant: select the circumstance with the smallest p-value and proceed
3. **Choose split:** for every possible way the selected circumstance can divide the sample into two subgroups, test the hypothesis of same mean outcome in the two resulting subgroups. Choose the splitting point with the smallest p-value
4. **Repeat :)**


# **Results**
```{r echo=T, message=FALSE, error=FALSE, warning=FALSE}
set.seed(123)

equality_data <- equality_data %>%
  mutate(train_index = sample(c("train", "test"), nrow(equality_data), replace=TRUE, prob=c(0.85, 0.15)))

train <- equality_data %>% filter(train_index=="train")
test <- equality_data %>% filter(train_index=="test")

formula <- income_k ~ Sex + Citizenship + Parents_Present + Adults_home + Father_CoB + Father_Cit + Mother_CoB + Mother_Cit + Father_Educ + Mother_Educ + Father_Occup_Stat + Mother_Occup_Stat + Father_Main_Occup + Mother_Main_Occup + Father_Manag + Mother_Manag + Tenancy

tree <- rpart(formula, data = train, cp=.001)

tree

rpart.plot(tree, box.palette="RdBu", nn=FALSE, type=2)

```
*Confusion Matrix
```{r echo=T, message=FALSE, error=FALSE, warning=FALSE}
test$prediction_tree <- predict(tree, newdata = test, type = c("matrix"))

class(test$prediction_tree)
levels(test$prediction_tree)

confusion <- confusionMatrix(test$income_k, test$prediction_tree, mode="everything")
confusion
```

```{r echo=T, message=FALSE, error=FALSE, warning=FALSE}
control <- trainControl(method = "repeatedcv", number = 10, repeats = 10, savePredictions = T, classProbs = T, summaryFunction = multiClassSummary)

tuning_grid <- expand.grid(cp = seq(0, 0.02, by= 0.005))
tuning_grid

tree_caret <- train(data = equality_data, formula, method = "rpart", trControl = control, tuneGrid = tuning_grid, metric = "Accuracy", na.action = na.pass)
tree_caret
```
# **Conclusion**
=======
# **Conclusion**

# **References**
>>>>>>> 8860f3d4dd72f45073c60dda2f3226dd505e53a1